Inference of Heating Properties from ``Hot'' Non-flaring Plasmas in Active Region Cores II. Nanoflare Trains

### 2016-08-10 hot_non-flaring_plasma_2/hot_non-flaring_plasma_2.tex
rewrite of section 3.2.2 and conclusions by PC, small edits

	\par One characteristic of the $\mathrm{EM}(T)$ distribution arising from low- and intermediate-frequency nanoflares is that $\mathrm{EM}(T)$ is dual-valued as a function of temperature \citep[see \autoref{fig:single_em}-\autoref{fig:ion_em} and][]{cargill_implications_1994}. For example, in the lower right panel of \autoref{fig:single_em}, $\mathrm{EM}\approx 10^{26.5}$ cm$^{-5}$ at $T = 10^{6.2}$ K and $10^{6.9}$ K. Given two observed temperatures (due say to specific emission lines), this suggests that predictions from the ratio of emission at these temperatures can be compared with data. This is of particular use where very limited spectral coverage is available either above or below $T_m$. Indeed for impulsive heating of any kind, the generic $\mathrm{EM}(T)$ curve remains similar (though not identical) as the magnitude of the heating changes. For flare-like energy release one would expect the hot emission to be much greater than the cold emission, with the opposite in the quiet Sun. For nanoflares, one might expect them to be of the same order.

@@

	\par An example of how this approach might work can be found in \citet{brosius_pervasive_2014}. Using observations of an active region from the \textit{EUNIS-13} sounding rocket, they found that the intensity ratio of Fe XIX (formed at $T\approx10^{6.95}$ K) to Fe XII (formed at $T\approx10^{6.2}$ K) is $\sim0.59$ inside the AR core as compared to $\sim0.076$ outside. They argue that this provides possible evidence for impulsive heating. \authorcomment1{These next few sentences need tightening up; next round of edits...have updated to be consistent with new choice for em ratio}As a proxy for this intensity ratio and an alternative to the hot emission measure slope $b$, we propose using an emission measure ratio, $\mathrm{EM}(T_{hot})/\mathrm{EM}(T_{cool})$. Unlike an emission measure slope, where a functional relationship is assumed over a range of $T$ in $\mathrm{EM}(T)$, a simple emission measure ratio gives information about the emission at only a single ``hot'' temperature relative to a single ``cool'' temperature. Using this ratio, we can assess the relative strength of $\mathrm{EM}(T)$ above and below $T_m$ and in doing so, assess the expected range of emission measure ratios consistent with nanoflare heating. We choose $T_{hot}=10^{6.95}$ K and $T_{cool}=10^{6.2}$ K, consistent with \citet{brosius_pervasive_2014} The ratio of other emission lines could also be used. 
	\par Such an emission measure ratio also provides a way to compare in a concise way every point in our multidimensional parameter space while acknowledging that we are reducing $8\times10^4$ s of loop evolution to a single value. The left panel of \autoref{fig:em_ratio_hist_single} shows the single-fluid case, where each individual histogram (denoted by linestyle and color) corresponds to a different type of heating function. This means, for example, that the solid blue histogram includes emission measure ratios for all values of $t_N$, but for only those cases where heating events are chosen from a power-law distribution of $\alpha=-1.5$ and $t_{N,i}=t_N$ for all $i$ (i.e. $\beta=0$). We see that the emission measure ratio is largely insensitive to $\alpha$ and is peaked sharply at $\sim0.75-1.0$; the distribution peaks at slightly higher values for the $\beta=1$ case ($\sim1.25$) though the overall shape of the distributions is the same. Note that the uniform heating results (whose median is denoted by the vertical purple line near $\sim0.5$) show emission measure ratios significantly less than those in the power-law cases, consistent with the reduced hot emission shown in \autoref{fig:el_em}. 
	\par The right panel of \autoref{fig:em_ratio_hist_single} shows these same emission measure ratios, but now separated by $t_N$. For example, the solid red histogram includes emission measure ratios for every type of heating function (i.e. uniform, all $\alpha$ and all $\beta$), but for only those runs where $t_N=3000$ s. Note that we choose to only show results for five values of $t_N$ for aesthetic purposes. For $t_N\le4000$ s, the results cluster just near 1.0, though the $t_N=3000$ s distribution is slightly bimodal, peaking strongly at $\sim0.75$ and more weakly at $\sim1.25$. Additionally, the $t_N=5000$ s case peaks slightly higher at $\sim1.5$ and does not ``pile up'' near 1.0 as the other values of $t_N$ do. In general, the distributions for all $t_N$ have a positive skew and are peaked in the range $\sim0.5-1.5$.
	\par Similarly, \autoref{fig:em_ratio_hist_el} shows the emission measure ratio for the case in which only the electrons are heated. Looking first at the left panel, we see that, as in \autoref{fig:em_ratio_hist_single}, the emission measure ratio is largely insensitive to $\alpha$, but is comparatively more narrow and falls off more quickly twoard higher emission measure ratios. The $\beta=0$ cases (for all $\alpha$) peak near $0.5$, with the $\beta=1$ cases all peaking at slightly higher values, but all less than 1. The right panel shows the same emission measure ratios now grouped by $t_N$. The $t_N=3000$ distribution has a very strong and narrow peak at $\approx0.5$ while all of the other distributions peak at $\sim1$. For $t_N\ge3000$ s, the location of the peak increases weakly with increasing $t_N$ though this is not the case for $t_N=1000,2000$ s. 
	\par Finally, we compute the same emission measure ratio for the case in which only the ions are heated. The left panel of \autoref{fig:em_ratio_hist_ion} shows histograms for each type of heating function. Similar to the electron  heating results, we find that the $\beta=0$ distributions peak at lower values compared to the $\beta=1$ distributions and that in the $\beta=0$ case, the distribution of emission measure ratios is relatively insensitive to $\alpha$. However, unlike the electron heating case, the distributions in all cases are much wider and peak at higher values, $\sim1.5$ for $\beta=0$ and $>2$ for $\beta=1$. Furthermore, in the $\beta=1$ case, the $\alpha=-2.5$ distribution peaks at lower values compared $\alpha=-1.5,-2.0$. The right panel of \autoref{fig:em_ratio_hist_ion} shows the same emission measure ratios now grouped by waiting time where again we only show five values of $t_N$. Compared to the results in the left panels of \autoref{fig:em_ratio_hist_single} and \autoref{fig:em_ratio_hist_el}, the distributions for each value of $t_N$ are much wider and the peak values show a stronger dependence on $t_N$. The range of peak values is also much larger, with the $t_N=1000$ s peaking near 0 and $t_N=5000$ s peaking at $\sim2$. As in \autoref{fig:em_ratio_hist_single}, we see that the $t_N=3000$ s distribution is bimodal. 
	\par\authorcomment2{These results suggest that emission measure ratios, for $T_{hot}=10^{6.95}$ K and $T_{cool}=10^{6.2}$ K, consistent with nanoflare heating fall approximately in the range $\sim0.5-1.5$. We find that the resulting distributions of ratios are largely independent of the power-law index $\alpha$ and only weakly-dependent on the heating frequency, $t_N$ (with the exception of the ion heating case). Furthermore, the distributions are also weakly dependent on the relationship between the waiting time and heating rate, $\beta$. In this way, a determination of more detailed nanoflare properties from observations may be difficult without a very large number of measurements and relatively small uncertainties in the atomic data. However, while the single-fluid and electron heating distributions are relatively similar (see \autoref{fig:em_ratio_hist_single} and \autoref{fig:em_ratio_hist_el}), the resulting ion heating distributions are much wider, peak at higher values, and show a stronger dependence on $\alpha$, $\beta$, and $t_N$. We may interpret this as the emission on the hot side of $\mathrm{EM}(T)$ being relatively insensitive to these parameter changes when only the ions are heated. In this way, the emission measure ratio probes mostly changes in the cool emission which we already know to be sensitive to both $\beta$ and $t_N$ \citep{cargill_active_2014}.}

@@

	\par Two possible ways to relate these results to present and future observations were discussed. First we showed that while below $T_m$ a power law representation of $\mathrm{EM}\sim T^a$ was quite robust, that was not the case above $T_m$. On the other hand the calculation of a ratio between the emission measure at a pair temperatures above and below $T_m$ held more promise. \authorcomment2{By using this ratio as an alternative to the emission measure slope, we do not assume $\mathrm{EM}(T)$ to have a particular functional form and we account for emission on the hot and cool side of $\mathrm{EM}(T)$, both of which contain signatures of the heating.}

@@

	\par\authorcomment2{While our emission measure ratio results show a relatively weak dependence on the particular properties of the heating (e.g. $\alpha,\beta,t_N$), we find that, for the electron heating, ion heating, and single-fluid cases, the peaks of the distributions all fall in the range $~0.5-1.5$. In this way, we find that the 0.59 emission measure ratio of \citet{brosius_pervasive_2014}, which they conclude to be a signature of nanoflares, is consistent with nanoflare heating for a range of different heating properties. A larger sample of observations in this wavelength range combined with the analysis presented here will help to constrain the heating properties of nanoflares.}
	\acknowledgments
	WTB was provided travel support to the Coronal Loops Workshop VII held in Cambridge, UK, at which a preliminary version of this work was presented, by NSF award number 1536094. This work was supported in part by the Big-Data Private-Cloud Research Cyberinfrastructure MRI-award funded by NSF under grant CNS-1338099 and by Rice University.
	\software{astroML \citep{vanderplas_introduction_2012}, IPython \citep{perez_ipython:_2007}, Jupyter notebook, matplotlib \citep{hunter_matplotlib:_2007}, NumPy, seaborn \citep{waskom_seaborn:_2016}, scipy \citep{van_der_walt_numpy_2011}}
	%Bibliography
	\bibliography{astrophys-abbrev.bib,references.bib}
	\bibliographystyle{aasjournal}
	%\bibliography{hot_non-flaring_plasma_2.bbl}
\end{document}

>>>>

		\par\citet{brosius_pervasive_2014} proposed another possible diagnostic for the hot non-flaring corona. Using observations of an active region from the \textit{EUNIS-13} sounding rocket, they found that the intensity ratio of Fe XIX (formed at $T\approx10^{6.95}$ K) to Fe XII (formed at $T\approx10^{6.2}$ K) is $\sim0.59$ inside the AR core as compared to $\sim0.076$ outside. They argue that this provides possible evidence for impulsive heating. This arises from the fact that the $\mathrm{EM}(T)$ distribution arising from low- and intermediate-frequency nanoflares is dual-valued as a function of temperature \citep[see \autoref{fig:single_em}-\autoref{fig:ion_em} and][]{cargill_implications_1994}: for example, in the lower right panel of \autoref{fig:single_em}, $\mathrm{EM}\approx 10^{26.5}$ cm$^{-5}$ at $T = 10^{6.2}$ K and $10^{6.9}$ K. This suggests that the approach of \citet{brosius_pervasive_2014} could be used for any pair of appropriate emission lines and is likely to be of particular use where very limited spectral coverage is available either above or below $T_m$. Indeed for impulsive heating of any kind, the generic $\mathrm{EM}(T)$ curve remains similar (though not identical) as the magnitude of the heating changes. For flare-like energy release one would expect the hot emission to be much greater than the cold emission, with the opposite in the quiet Sun.

	\par We define a general emission measure ratio, $\mathrm{EM}(T_{hot})/\mathrm{EM}(T_{cool})$ and for this paper consider $T_{hot}=10^{6.95}$ K and $T_{cool}=10^{6.2}$ K, consistent with \citet{brosius_pervasive_2014}. This also provides a way to compare in a concise way every point in our multidimensional parameter space while acknowledging that we are reducing $8\times10^4$ s of loop evolution to a single value. This ratio is shown in \autoref{fig:em_ratio_hist_single}, \autoref{fig:em_ratio_hist_el}, and \autoref{fig:em_ratio_hist_ion} for the single-fluid model, electron and ion heating, respectively. In the left-hand panels each individual histogram (denoted by linestyle and color) corresponds to a different type of heating function. This means, for example, that the solid blue histogram includes emission measure ratios for all values of $t_N$, but for only those cases where heating events are chosen from a power-law distribution of $\alpha=-1.5$ and $t_{N,i}=t_N$ for all $i$ (i.e. $\beta=0$). The right panel of \autoref{fig:em_ratio_hist_single} shows these same emission measure ratios, but now categorized by $t_N$. For example, the solid red histogram includes emission measure ratios for every type of heating function (i.e. uniform, all $\alpha$ and all $\beta$), but for only those runs where $t_N=3000$ s. Note that we choose to only show results for five values of $t_N$ for aesthetic purposes.

	\par Considering first the left hand panels, in the single fluid model the ratio is largely insensitive to $\alpha$ and is peaked sharply at $\sim0.75-1.0$ although the distribution peaks at slightly higher values for the $\beta=1$ case ($\sim1.25$). Note that the uniform heating results (whose median is denoted by the vertical purple line near $\sim0.5$) show emission measure ratios significantly less than those in the power-law cases, consistent with the reduced hot emission shown in \autoref{fig:single_em}. For electron heating the ratio is again insensitive to $\alpha$, but is narrower and falls off more quickly toward higher emission measure ratios. The $\beta=0$ cases (for all $\alpha$) peak near $0.5$, with the $\beta=1$ cases all peaking at slightly higher values, but all less than 1. For ion heating the $\beta=0$ distributions peak at lower values compared to those with $\beta=1$. Again the results are relatively insensitive to $\alpha$. However, compared to the electron heating case, all of the distributions are much wider and peak at higher values, $\sim1.5$ for $\beta=0$ and $>2$ for $\beta=1$. Furthermore, in the $\beta=1$ case, the $\alpha=-2.5$ distribution peaks at lower values compared $\alpha=-1.5,-2.0$.

	\par Turning to the right hand panels, in the single-fluid model for $t_N\le4000$ s, the results cluster just near 1.0, though the $t_N=3000$ s distribution is slightly bimodal, peaking strongly at $\sim0.75$ and more weakly at $\sim1.25$. Additionally, the $t_N=5000$ s case peaks slightly higher at $\sim1.5$ and does not ``pile up'' near 1.0 as the other values of $t_N$ do. For electron herating, the $t_N=3000$ distribution has a very strong and narrow peak at $\approx0.5$ while all of the other distributions peak at $\sim1$. For $t_N\ge3000$ s, the location of the peak increases weakly with increasing $t_N$ though this is not the case for $t_N=1000,2000$ s. Finally, for ion heating the distributions for each value of $t_N$ are much wider than for a single fluid and electron heating and the peak values show a stronger dependence on $t_N$. The range of peak values is also much larger, with the $t_N=1000$ s case peaking near 0 and the $t_N=5000$ s case peaking at $\sim2$. As in \autoref{fig:em_ratio_hist_single}, we see that the $t_N=3000$ s distribution is bimodal. In general, the distributions grouped by $t_N$ for all three heating types have a positive skew and are peaked in the range $\sim0.5-1.5$ except for some extreme case in the ion heating scenario.

@@

	\par These results suggest that the emission measure ratio is generally in the range $\sim0.5-1.5$, with some higher values. Given the uncertainties in the atomic physics in particular, this seems to support the conclusion of the presence of nanoflare heating by \citet{brosius_pervasive_2014}. Thus the calculation of such ratios from limited data has the potential to be a powerful diagnostic of the existence of nanoflare heating. Whether one can say more about the precise form of nanoflare heating is less clear. In these results the ratios are largely independent of the power-law index $\alpha$ and only weakly-dependent on the heating frequency, $t_N$. Furthermore, the distributions are also weakly dependent on the relationship between the waiting time and heating rate, $\beta$. The exception is the ion heating case where the distributions are much wider, peak at higher values, and show a stronger dependence on $\alpha$, $\beta$, and $t_N$. The problem with drawing conclusions from the details of these results lies in the numerous uncertainties in any data, especially concerning atomic physics.

@@

	\par Two possible ways to relate these results to present and future observations were discussed. First we showed that while below $T_m$ a power law representation of $\mathrm{EM}\sim T^a$ was quite robust, that was not the case above $T_m$. On the other hand the calculation of a ratio between the emission measure at a pair of temperatures above and below $T_m$ held more promise. For a wide range of parameters the ratio of the emission measure at two temperatures, $T_{hot}>T_m$ at $10^{6.9}$ and $T_{cool}<T_m$ at $10^{6.2}$ was of order unity, consistent with \citet{brosius_pervasive_2014} within various errors. Further observational evidence of such ratios would, in our view, provide a very strong case for the presence of nanoflare heating. On the other hand, given the uncertainties in atomic physics and emission emasure analysis, such a comparison seems unlikely to be able to shed more detailed light on the details of the actual heating process. 

@@

	\par It is clear from this and the preceeding paper that characterization of this ``hot'' component is extremely challenging for a multiplicity of reasons. Progress is likely to come from the good spectral coverage of the MaGIXS instrument, and in particular from a space-based successor which could confirm beyond doubt the presence of the hot component. On the other hand, it seems doubtful that information regarding the plasma physics of the heating process can come from such instrumentation. Information from high-energy instruments would be then desirable, but the energies of interest ($\approx1$ keV) are highly challenging. In any event, complete wavelength coverage seems essential, something that we noted earlier stellar astronomers have had for decades.

	\acknowledgments
	%
	WTB was provided travel support to the Coronal Loops Workshop VII held in Cambridge, UK, at which a preliminary version of this work was presented, by NSF award number 1536094. This work was supported in part by the Big-Data Private-Cloud Research Cyberinfrastructure MRI-award funded by NSF under grant CNS-1338099 and by Rice University.
	%
	\software{astroML \citep{vanderplas_introduction_2012}, IPython \citep{perez_ipython:_2007}, Jupyter notebook, matplotlib \citep{hunter_matplotlib:_2007}, NumPy, seaborn \citep{waskom_seaborn:_2016}, scipy \citep{van_der_walt_numpy_2011}}
	%
	%
	%Bibliography
	\bibliography{astrophys-abbrev.bib,references.bib}
	\bibliographystyle{aasjournal}
	%\bibliography{hot_non-flaring_plasma_2.bbl}
\end{document}
